\section{Method}

\begin{frame}{Method}
  \begin{itemize}
    \item  \textbf{Feature localization}: Query2Label: A Simple Transformer Way to Multi-Label
Classification (Q2L)
    \item \textbf{Sparse label annotation}: Multi-Label Learning from Single Positive Labels (MLSPL)
  \end{itemize}
\end{frame}

\begin{frame}{Query2Label: Core Idea}
  \begin{itemize}
    \item Treat each label as a \emph{learnable query} in a Transformer decoder.
    \item Decoder cross‑attends to backbone feature map.
    \item Produces class‑specific feature vectors.
  \end{itemize}
\end{frame}
%-------------------------------------------------

\begin{frame}{Q2L Architecture}
  \begin{itemize}
    \item Stage 1: CNN/ViT backbone extracts spatial features.
    \item Stage 2: Multi‑layer Transformer decoder with $K$ queries.
    \item Linear head → $K$ sigmoid outputs (one per label).
    \item Loss: Asymmetric Focal (AF) to mitigate imbalance.
  \end{itemize}
\end{frame}
%-------------------------------------------------

\begin{frame}{Q2L Training Setup}
  \begin{itemize}
    \item 80 epochs, AdamW, RandAugment, EMA.
    \item Backbones: ResNet‑101, TResNet‑L, Swin‑L, CvT‑w24.
    \item Memory constraints encountered on 12 GB GPUs.
  \end{itemize}
\end{frame}
%-------------------------------------------------

\begin{frame}{MLSPL: Learning from One Positive Label}
  \begin{itemize}
    \item Extreme weak supervision: exactly \textbf{one} positive per image.
    \item Objective: recover missing positives during training.
  \end{itemize}
\end{frame}
%-------------------------------------------------

\begin{frame}{ROLE: Regularised Online Label Estimation}
  \begin{itemize}
    \item Maintain soft estimates $\hat{y}$ for unobserved labels.
    \item Jointly optimise classifier \& label estimator.
    \item Batch‑level regulariser keeps expected positives $\approx \kappa$.
  \end{itemize}
\end{frame}
%-------------------------------------------------

\begin{frame}{Dataset: MS--COCO 2014}
  \begin{itemize}
    \item 82 k train / 40 k val images, 80 object categories.
    \item Avg. 2.9 labels / image.
    \item Single‑positive variant generated for MLSPL.
  \end{itemize}
\end{frame}
%-------------------------------------------------
